---
title: "Final Project"
output: html_notebook
Authors: Danielle Buhr and Gerardo Sandoval
---

 
```{r}

library(ggplot2) # for graphs
library(rjags)
library(dplyr)
#library(fBasics)
#library(forecast)

```


## Load Data
```{r}
spy_path = paste(getwd(), '/SPY.csv', sep='')

spy_data = read.csv(spy_path)
head(spy_data)
#View(spy_data)#

```


## Variables
```{r}
#### Create Change Variables ####
spy_data = spy_data %>%
  dplyr::mutate(change1 = (Adj.Close - lag(Adj.Close, n=1))/ lag(Adj.Close, n=1),
                change3 = (Adj.Close - lag(Adj.Close, n=3))/ lag(Adj.Close, n=3),
                change5 = (Adj.Close - lag(Adj.Close, n=5))/ lag(Adj.Close, n=5),
                change10 = (Adj.Close - lag(Adj.Close, n=10))/ lag(Adj.Close, n=10),
                change0 = (Close - Open) / Open
                )
hist(spy_data$change3, breaks=50)

```



*G.S.*: I was thinking of analyzing the absolute values to make the logistic regression work a bit easier

```{r}
abs_data = spy_data %>%
  dplyr::mutate(change1 = abs((Adj.Close - lag(Adj.Close, n=1))/ lag(Adj.Close, n=1)),
                change3 = abs((Adj.Close - lag(Adj.Close, n=3))/ lag(Adj.Close, n=3)),
                change5 = abs((Adj.Close - lag(Adj.Close, n=5))/ lag(Adj.Close, n=5)),
                change10 = abs((Adj.Close - lag(Adj.Close, n=10))/ lag(Adj.Close, n=10))
                )


hist(abs_data$change3, breaks=50)

mean(na.omit(abs_data$change3))

```


#### Volume Variables
```{r}
# Function to normalize volume
norm_volume = function(data, new_max, new_min){
  return(data-min(data)) / (max(data)-min(data)) * (new_max-new_min) + new_min
}

### Normalized Volume ###
spy_data$Norm_Volume = norm_volume(spy_data$Volume, 3, 1)


```

### Advance Decline Variables
```{r}

# mutate is from dplyr
spy_data = spy_data %>%
  mutate(Volume_Dir = (change0/abs(change0))*Norm_Volume,
         Adv_Dec = cumsum(ifelse(is.na(Volume_Dir), 0, Volume_Dir)),
         AD_Change1 = (Adv_Dec - lag(Adv_Dec)) / lag(Adv_Dec),
         AD_Change3 = (Adv_Dec - lag(Adv_Dec,n=3)) / lag(Adv_Dec,n=3),
         AD_Change5 = (Adv_Dec - lag(Adv_Dec,n=5)) / lag(Adv_Dec,n=5)
         )
plot(spy_data$Adv_Dec, type='l', xlim=c(1000,5000))
plot(spy_data$AD_Change1, type='l', xlim=c(1000,5000), ylim=c(-1,1))

sd(spy_data$change1)
sd(spy_data$change3)
```


#### Response Variable
*G.S.*: We want to predict what the change will be from today over the next 5 days. So I shifted the change5 column to create a response column
```{r}

spy_data$Response = lead(spy_data$change5, n=5)

check_values = subset(spy_data, select = c(Response, change5))
head(check_values, n=10)

```


## JAG Structure
*DB:* are the params correct? I think we're just traking betas

```{r}
spy_data = na.omit(spy_data)
Y = spy_data$Response
X = subset(spy_data, select=c(Norm_Volume, change1, change3, Adv_Dec))
X1 = spy_data$Norm_Volume
X2 = spy_data$change1
X3 = spy_data$change3
X4 = spy_data$Adv_Dec

n = length(Y)
p = ncol(X)
data = list(Y=Y, X1=X1, X2=X2, X3=X3, X4=X4, n=n)#, p=p)
params = c('alpha', 'beta1', 'beta2', 'beta3', 'beta4') 
names = c('alpha',"Norm_Volume", "change1", 'change3', 'Adv_Dec')

nBurn = 10000
nChains = 2
nSave = 4000
nThin = 5
nIter = ceiling((nSave*nThin)/nChains)

```


## JAGS Model (1.)

```{r}
model_string1 = textConnection("
  model{
  # Likelihood
  for(i in 1:n){
    Y[i] ~ dnorm(alpha + X1[i]*beta1+ X2[i]*beta2+ X3[i]*beta3+ X4[i]*beta4, tau1)
  }
  # Priors
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 1/0.012)
  beta3 ~ dnorm(0, 1/0.019)
  beta4 ~ dnorm(0, 0.001)
  tau1 ~ dgamma(0.1, 0.1) 
  alpha ~ dnorm(0, 0.001)
  }")

model1 = jags.model(model_string1, data=data, n.chains=nChains, quiet=T)
update(model1, burn=nBurn, progress.bar='text')
samples1 = coda.samples(model1, variable.names=params, thin=nThin, n.iter=nIter, progress.bar='text')


```


### RESULTS (1.)
```{r}
plot(samples1)
round(effectiveSize(samples1), 1)
sum1 = summary(samples1)
rownames(sum1$statistics) = names
rownames(sum1$quantiles) = names
sum1

```


## JAGS Model (2.)
```{r}
model_string2 = textConnection("
  model{
  # Likelihood
  for(i in 1:n){
    Y[i] ~ dnorm(alpha + X1[i]*beta1+ X2[i]*beta2+ X3[i]*beta3+ X4[i]*beta4, tau1)
  }
  # Priors
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 0.001)
  beta3 ~ dnorm(0, 0.001)
  beta4 ~ dnorm(0, 0.001)
  tau1 ~ dgamma(0.1, 0.1) 
  alpha ~ dnorm(0, 0.001)
  }")

model2 = jags.model(model_string2, data=data, n.chains=nChains, quiet=T)
update(model2, burn=nBurn, progress.bar='text')
samples2 = coda.samples(model2, variable.names=params, thin=nThin, n.iter=nIter, progress.bar='text')
```

### RESULTS (2.)
D.B: edits made
```{r}
plot(samples2)
round(effectiveSize(samples2), 1)
sum2 = summary(samples2)
rownames(sum2$statistics) = names # pointed to sum1 instead of 2
rownames(sum2$quantiles) = names
sum2

```
## JAGS Model (3.)
DB edits
```{r}
data3 = list(Y=Y, X1=X1, X2=X2, X3=X3, X4=X4, n=n)
params3 = c('alpha', 'beta')
```


```{r}
model_string3 = textConnection("
  model{
  # Likelihood
  for(i in 1:n){
    Y[i] ~ dnorm(mu[i],taue)
    mu[i] ~ dnorm(alpha + X1[i]*beta[1]+ X2[i]*beta[2]+ X3[i]*beta[3]+ X4[i]*beta[4], taue)
    
  }
  
 
  # Priors
  beta[1] ~ dnorm(0, 0.001)
  beta[2] ~ dnorm(0, 0.001)
  beta[3] ~ dnorm(0, 0.001)
  beta[4] ~ dnorm(0, 0.001)
  taue ~ dgamma(0.1, 0.1) 
  alpha ~ dnorm(0, 0.001)
  }")

model3 = jags.model(model_string3, data=data3, n.chains=nChains, quiet=T)
update(model3, burn=nBurn, progress.bar='text')
samples3 = coda.samples(model3, variable.names=params3, thin=nThin, n.iter=nIter, progress.bar='text')
```

### RESULTS (3.)
```{r}
plot(samples3)
round(effectiveSize(samples3), 1)
sum3 = summary(samples2)
rownames(sum3$statistics) = names
rownames(sum3$quantiles) = names
sum3

```



### Notes
```{r}
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 0.001)
  beta3 ~ dnorm(0, 0.001)
  beta4 ~ dnorm(0, 0.001)
  
#### Part of Model ####
for(j in 1:n){
tau[1] ~ dgamma(0.1,0.1)
}

for(j in 2:n){
rho ~ runif(-1, 1)
}


#### Volume ####
volume = spy_data$Volume
intraday_change = (spy_data$Close - spy_data$Open) / spy_data$Open

plot(cumsum(volume*(intraday_change/abs(intraday_change))))

#
 Y[i] ~ dnorm(beta[i,1]+beta[i,2]+beta[i,3]+beta[i,4]+beta[i,5]+beta[i,6])/6+bias3,tau3)
  for(j in 1:n){
    tau[1] ~ dgamma(0.1,0.1)
  }
  
  for(j in 2:n){
  rho ~ dunif(-1, 1)
  }
  # Priors
  alpha ~ dnorm(0, 0.001)
  for (j in 1:p){
    beta[j]~ dnorm(0, 0.001)
  }
  tau1 ~ dgamma(0.1, 0.1)
```
